---
title: "RDA_Codes_formatted"
author: "Thomas Van Zuiden"
date: "October 24, 2016"
output: html_document
---

##Loading libraries
``` {r}
library(ade4)
library(vegan)
library(MASS)
install.packages("packfor", repos="http://R-Forge.R-project.org")
library(packfor)
```

##Load data and log-transform environmental variables that violate statistical assumptions
``` {r}
setwd("C:\\Users\\adminuser\\Desktop\\EVERYTHING THOMAS\\Chapter 2")
dat<-read.csv("rda.csv")
dat<-na.omit(dat)
nrow(dat)
class(dat)
dat <- as.data.frame(dat)
colnames(dat)
```

###Write a loop to log-transform the necessary variables
```{r}
lakedat <- cbind(dat[4:9])
for (i in 1:6){
    lakedat[i] <- log10(lakedat[i]+1) 
}
climdat <- cbind(dat[10:19])
fishdat <- cbind(dat[20:29])
envdat <- data.frame(lakedat, climdat)
```
## Hellinger-transform species data and standardize environmental variables
``` {r}
spe.hel=decostand(fishdat, "hellinger") ### Hellinger transform species data
env.std=decostand(envdat, "standardize") ### Standardize environmental data
```

## Trying out an RDA with all explanatory variables in the data
### It's going to be messy
``` {r}
fishRDA=rda(spe.hel ~ ., data=env.std)
R2adjusted=RsquareAdj(fishRDA)$adj.r.squared
R2adjusted
plot(fishRDA)
text(fishRDA, display = "species", cex=0.8, col="red")
```

## Cleaning the RDA to eliminate redundant constraints
``` {r}
#Forward selection will indicate important explanatory variables
forward.sel(spe.hel, env.std)

#Recreating a cleaner RDA with only important predictors of fish communities
fishRDA=rda(spe.hel ~Surface_Area_ha+Secchi_Depth_m+July_0012_temp+July_0012_pre+DOC_mgL+P_ugL, data=env.std)

#Global adjusted R2
R2adjusted=RsquareAdj(fishRDA)$adj.r.squared
R2adjusted

#Assessing multicollinearity between explanatory varables
vif.cca(fishRDA) ## values over 10 indicate redundant constraints

#Plot RDA with only significant variables
plot(fishRDA)
text(fishRDA, display = "species", cex=0.8, col="red")
anova.cca(fishRDA, step=1000)
```



